---
title: "temp + otolith data"
author: "Sarah Willington"
date: "2026-01-28"
output: html_document
---

script for all data manipulation for otolith paper. Chapter 3 of my PhD

1. manipulate temp data so that annual means are from oct - sep and summer means are from oct - march

2. merge otolith and temp datasets

3. sub sample bmw, pw and sfh data 

4. add maximum age col

5. calculate temp variables for model, scale and set up data



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


```

```{r load libraries}
library(tidyverse)
library(here)
library(patchwork)
library(colorspace)

library(rlang)


```

```{r load otolith data}
source("01_code/10_importing_otolith_data_all.R")

```

```{r load temp data}
project_dir <- here::here()

hadisst_path_monthly <- file.path(project_dir, "0_data", "01_raw_data", "HadISST_all_sites_monthly_mean.csv")
hadisst_monthly <- read_csv(hadisst_path_monthly) |> 
  rename(fishyear = year) |> 
  filter(site!="PPB" & site!="WP")

```
 
### 1. manipulate temp data

```{r temp data wrangling HadISST}
# Oct - Dec temp data
hadisst_oct <- hadisst_monthly |> 
  filter(month>9) |>      
  arrange(fishyear, .by_group = TRUE)

#Jan - Sep temp data
hadisst_sep <- hadisst_monthly |> 
  filter(month<10) |>       
  arrange(fishyear, .by_group = TRUE) |>  
  mutate(fishyear_neg = fishyear - 1)  |>  #create new col, of the year previous to match up with the oct-dec data, so that when take the mean temp in that year you are taking the correct month-year combo
  select(-fishyear) |> 
  rename(fishyear = fishyear_neg)

hadisst_annual <- bind_rows(hadisst_oct, hadisst_sep) |> 
  filter(fishyear>1904 & fishyear<2025) |> 
  group_by(site, fishyear) |> 
  summarise(sst_annual = mean(sst_mean), .groups = "drop") |> 
  arrange(site, fishyear) |>                      # ensure correct order
  group_by(site) |>                               # lag within site
  mutate(sst_annual_lag = dplyr::lag(sst_annual)) |> 
  unite(siteyear, c(site, fishyear), sep = "_", remove = TRUE) |> 
  ungroup()


## Mean summer temp
#Oct - Dec temp data
hadisst_oct <- hadisst_monthly |> 
  filter(month>9)|>        
  arrange(fishyear, .by_group = TRUE)

#Jan - Mar temp data
hadisst_jan <- hadisst_monthly |> 
  filter(month<4)|>        
  arrange(fishyear, .by_group = TRUE) |> 
  mutate(fishyear_neg = fishyear - 1) |> #create new col, of the year previous to match up with the oct-dec data, so that when take the mean temp in that year you are taking the correct month-year combo
  select(-fishyear) |> 
  rename(fishyear = fishyear_neg)

hadisst_summer <- bind_rows(hadisst_oct, hadisst_jan)|> 
  filter(fishyear>1904 & fishyear<2025) |> 
  group_by(site, fishyear) |> 
  summarise(sst_summer = mean(sst_mean), .groups = "drop") |> 
  arrange(site, fishyear) |>                      # ensure correct order
  group_by(site) |>                               # lag within site
  mutate(sst_summer_lag = dplyr::lag(sst_summer)) |> 
  unite(siteyear, c(site, fishyear), sep = "_", remove = TRUE) |> 
  ungroup()

```

### 2. merge otolith and temp datasets

```{r joining oto and temp data}
oto_incs <- oto_incs %>% 
  unite(siteyear, c(site, fishyear), sep = "_", remove = FALSE) |>  
  left_join(hadisst_annual, by = "siteyear") |> 
  left_join(hadisst_summer, by = "siteyear")

sfh <- sfh %>% 
  unite(siteyear, c(site, fishyear), sep = "_", remove = FALSE) |> 
  left_join(hadisst_annual, by = "siteyear") |> 
  left_join(hadisst_summer, by = "siteyear")

wrasse <- wrasse %>% 
  unite(siteyear, c(site, fishyear), sep = "_", remove = FALSE) |>  
  left_join(hadisst_annual, by = "siteyear") |> 
  left_join(hadisst_summer, by = "siteyear") |> 
  mutate(monthcap = as.character(monthcap))

morwong <- morwong %>% 
  unite(siteyear, c(site, fishyear), sep = "_", remove = FALSE) |>  
  left_join(hadisst_annual, by = "siteyear") |> 
  left_join(hadisst_summer, by = "siteyear")|> 
  mutate(monthcap = as.character(monthcap)) 
# 
# write_csv(oto_incs, "0_data/02_cleaned_data/oto_incs_temperature.csv")
# write_csv(sfh, "0_data/02_cleaned_data/sfh_temperature.csv")
# write_csv(wrasse, "0_data/02_cleaned_data/wrasse_temperature.csv")
# write_csv(morwong, "0_data/02_cleaned_data/morwong_temperature.csv")

oto_incs_all <- bind_rows(oto_incs, sfh, wrasse, morwong)
# write_csv(oto_incs_all, "0_data/02_cleaned_data/oto_incs_full_df_temperature.csv")

```

### 3. sub sample


```{r subsample function}
 subsample_fish <- function(df,
                            fish_id_col   = "FishID",
                            age_col       = "age",
                            site_col      = "site",
                            n_per_group   = 20,
                            seed          = 1,
                            #breaks        = c(2, 5, 10, 15, 20, 25, 35, 50, 70, 90, 110),
                            breaks        = c(2, 10, 20, 35, 50, 75, 100),
                            right         = FALSE,
                            include_lowest = TRUE) {
   
   
   id_sym   <- rlang::sym(fish_id_col)
   age_sym  <- rlang::sym(age_col)
   site_sym <- rlang::sym(site_col)
   
   # Ensure breaks are sorted and unique
   breaks <- sort(unique(breaks))
   
   # Create labels 
   make_labels <- function(b) {
     labs <- character(length(b))
     for (i in seq_along(b)) {
       if (i < length(b)) {
         labs[i] <- paste0(b[i], "-", b[i + 1] - 1)
       } else {
         labs[i] <- paste0(b[i], "+")
       }
     }
     labs
   }
   labels <- make_labels(breaks)
   
   set.seed(seed)
   
   # 1) Collapse to unique fish: max age + a single site group per FishID
   fish_level <- df |>
     dplyr::group_by(!!id_sym) |>
     dplyr::summarise(
       max_age = max(!!age_sym, na.rm = TRUE),
       site_group = dplyr::first(!!site_sym),
       .groups = "drop"
     )
   
   # 2) Assign age bins based on max age
   # Using cut() with breaks + Inf, so the final bin is 110+
   fish_level <- fish_level |>
     dplyr::mutate(
       age_bin = cut(
         max_age,
         breaks = c(breaks, Inf),
         labels = labels,
         right = right,
         include.lowest = include_lowest
       )
     )
   
   # 3) Sample FishIDs within each (site_group x age_bin)
   sampled_ids <- fish_level |>
     dplyr::filter(!is.na(age_bin)) |>
     dplyr::group_by(site_group, age_bin) |>
     dplyr::summarise(
       FishID_sampled = list(sample(
         x = !!id_sym,
         size = min(n_per_group, dplyr::n()),
         replace = FALSE
       )),
       n_available = dplyr::n(),
       n_sampled = min(n_per_group, dplyr::n()),
       .groups = "drop"
     ) |>
     tidyr::unnest(FishID_sampled) |>
     dplyr::pull(FishID_sampled) |>
     unique()
   
   # 4) Return all rows for the sampled FishIDs
   df_sub <- df |>
     dplyr::filter(!!id_sym %in% sampled_ids)
   
   # Attach a little metadata as attributes (handy for debugging)
   attr(df_sub, "seed") <- seed
   attr(df_sub, "n_per_group") <- n_per_group
   attr(df_sub, "age_breaks") <- breaks
   df_sub
 }
```

```{r run subsample function}
# run function
 bmw_sub <- subsample_fish(
   df = morwong,
   fish_id_col = "FishID",
   age_col = "age",
   site_col = "site",
   n_per_group = 10,
   seed = 123
 )
 
 pw_sub <- subsample_fish(
   df = wrasse,
   fish_id_col = "FishID",
   age_col = "age",
   site_col = "site",
   n_per_group = 15,
   seed = 123
 )
 
 sfh_sub <- subsample_fish(
   df = sfh,
   fish_id_col = "FishID",
   age_col = "age",
   site_col = "site",
   n_per_group = 30,
   seed = 123
 )
 
 # Check how many unique FishIDs you got per group:
 
 pw_n <- pw_sub |> distinct(FishID, .keep_all = TRUE) 
 bmw_n <- bmw_sub |> distinct(FishID, .keep_all = TRUE) 
 sfh_n <- sfh_sub |> distinct(FishID, .keep_all = TRUE) 
 
 summary <- bmw_sub |> 
   group_by(site, FishID) |> 
   # summarise(max_age = max(age), .groups="drop") |> 
   # mutate(age_bin = cut(max_age, breaks = c(2,5,10,15,20,25,35,50,70,90,110, Inf),
   #                      labels = c("2-4","5-9","10-14","15-19","20-24","25-34", "35-49","50-69","70-89","90-109","110+"),
   #                      right = FALSE, include.lowest = TRUE)) |> 
   summarise(max_age = max(age), .groups="drop") |> 
   mutate(age_bin = cut(max_age, breaks = c(2, 10, 20, 35, 50, 75, 100, Inf),
                        labels = c("2-9","10-19","20-34","35-49","50-74", "75-99","100+"),
                        right = FALSE, include.lowest = TRUE)) |> 
   count(site, age_bin)

 # write_csv(bmw_sub, "0_data/02_cleaned_data/morwong_subsampled2_temperature.csv")
 # write_csv(pw_sub, "0_data/02_cleaned_data/wrasse_subsampled2_temperature.csv")
 # write_csv(sfh_sub, "0_data/02_cleaned_data/sfh_subsampled2_temperature.csv")
```

### 4. Add max age 


```{r add max age}

max_age <- read_csv(file.path(project_dir, "0_data", "01_raw_data", "species_max_age.csv"))
max_age <- max_age[1:2]

d <- bind_rows(bmw_sub, pw_sub, sfh_sub, oto_incs)
d_n <- d |> distinct(FishID, .keep_all = TRUE) 

d <- d |>  
  left_join(max_age, by = "species") 

```

### 5. add variables 


```{r}
hadisst_summer <- hadisst_summer |> 
  separate(siteyear, c("site", "year"), sep="_")

# get long term annual mean per site
sst_summer <- hadisst_summer |> 
  filter(year > 1923) |> 
  group_by(site) |> 
  mutate(sst_space = mean(sst_summer), 
         sst_space_lag = mean(sst_summer_lag)) |> 
  distinct(site, sst_space, sst_space_lag)

d <- d |> left_join(sst_summer, by = "site")

d <- d |>
  mutate(
    FishID = as.factor(FishID),
    cohort = as.factor(cohort),
    species = as.factor(species),
    log_increment = log(increment),
    log_age = log(age),
    sst_time = sst_summer - sst_space,
    species_cohort = interaction(species, cohort), 
    site_cohort = interaction(site, cohort),
    species_site_cohort = interaction(species, site, cohort)
  ) |>
  # now scale the temperatures
  mutate(
    sst_space_sc = (sst_space - mean(sst_space)) / sd(sst_space),
    sst_time_sc = (sst_time - mean(sst_time)) / sd(sst_time)
  ) 

#write_csv(d,  "0_data/02_cleaned_data/otolith_analysis.csv")
```

