---
title: "18_linear_model_current"
author: "Sarah Willington"
date: "2026-03-19"
output: html_document
---
Main analysis for Willington et al. 2026 otolith paper 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


```

```{r}
library(bbmle)
library(tidyverse)
library(here)
library(patchwork)
library(ggeffects) 
library(tidylog)
library(sdmTMB)
library(modelr)
```

### scaling and setting up variables

```{r scale variables}

project_dir <- here::here()

d <- read_csv(file.path(project_dir, "0_data", "02_cleaned_data", "otolith_analysis.csv"))

d <- d |>
  mutate(
    FishID = as.factor(FishID),
    cohort = as.factor(cohort),
    species = as.factor(species),
    species_cohort = as.factor(species_cohort),
    species_site_cohort = as.factor(species_site_cohort), 
    site_cohort = as.factor(site_cohort)
  ) 

d20 <- d |> 
  filter(age<=20) |> 
  mutate(age_rel = if_else(maximum_age > 20, age / 20, age / maximum_age),
         age_sc = (age - mean(age)) / sd(age),
         log_age_sc = (log_age - mean(log_age)) / sd(log_age),
         log_increment_sc = (log_increment - mean(log_increment)) / sd(log_increment),
         .by = species) |> 
  mutate(log_age_rel = log(age_rel))

d30 <- d |> 
  filter(age<=30) |> 
  mutate(age_rel = if_else(maximum_age > 30, age / 30, age / maximum_age),
         age_sc = (age - mean(age)) / sd(age),
         log_age_sc = (log_age - mean(log_age)) / sd(log_age),
         log_increment_sc = (log_increment - mean(log_increment)) / sd(log_increment),
         .by = species)|> 
  mutate(log_age_rel = log(age_rel))

dfull <- d |> 
  mutate(age_rel = age/maximum_age,
         age_sc = (age - mean(age)) / sd(age),
         log_age_sc = (log_age - mean(log_age)) / sd(log_age),
         log_increment_sc = (log_increment - mean(log_increment)) / sd(log_increment),
         .by = species)|> 
  mutate(log_age_rel = log(age_rel))

```

### fit models
#### maximum age 20 yrs

```{r test RE m20}

#base model
m20 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# random age slope per species - allowing species to have different growth ~ age relationship
m20a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# cohort nested in species - fish will have a different cohort experience depending on species
m22 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | species_cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m22a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | species_cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# cohort nested in site - fish will have a different cohort experience depending on where they are spatially
m23 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m23a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | site_cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# cohort nested in species in site - fish will have a different cohort experience depending on species and where they are spatially
m24 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | species_site_cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m24a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | species_site_cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

AICctab(m20, m20a, m22, m22a, m23, m23a, m24, m24a, base=TRUE, logLik=TRUE)
AICctab(m20, m20a, base=TRUE, logLik=TRUE)
AICctab(m20, m22,m23, m24, base=TRUE, logLik=TRUE)


```


```{r model checks m20}
qqnorm(residuals(m23))
qqline(residuals(m23))

tidy(m23)
sanity(m23)
summary(m23)

tidy(m23, effects = "ran_pars")

t <- tidy(m23, effects = "fixed") |>
  mutate(stat = estimate / std.error)

# Plot residuals vs age
d20$res <- residuals(m23)

ggplot(d20, aes(log_age_rel, res)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "tomato")+
  ggtitle("max age 20")


```

#### maximum age 30 yrs

```{r test RE m30}

#base model
m30 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# random age slope per species - allowing species to have different growth ~ age relationship
m30a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# cohort nested in species - fish will have a different cohort experience depending on species
m32 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | species_cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m32a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | species_cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# cohort nested in site - fish will have a different cohort experience depending on where they are spatially
m33 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m33a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | site_cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

# cohort nested in species in site - fish will have a different cohort experience depending on species and where they are spatially
m34 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | species_site_cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m34a <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (log_age_rel | species) + (1 | species_site_cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

AICctab(m30, m30a, m32, m32a, m33, m33a, m34, m34a, base=TRUE, logLik=TRUE)
AICctab(m30, m30a, base=TRUE, logLik=TRUE)
AICctab(m30, m32,m33, m34, base=TRUE, logLik=TRUE)

```


```{r model checks m30}
qqnorm(residuals(m33))
qqline(residuals(m33))

tidy(m33)
sanity(m33)
summary(m33)

tidy(m33, effects = "ran_pars")

t30 <- tidy(m33, effects = "fixed") |>
  mutate(stat = estimate / std.error)

# Plot residuals vs age
d30$res <- residuals(m33)

ggplot(d30, aes(log_age_rel, res)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "tomato")+
  ggtitle("max age 30")

ggplot(d20, aes(log_age_rel, res)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "tomato")+
  ggtitle("max age 20")

mfull <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = dfull, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

dfull$res <- residuals(mfull)

ggplot(dfull, aes(log_age_rel, res)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "tomato")+
  ggtitle("full max age")
```

The residuals look best in the maximum age capped at 20 years data, but both 20 and 30 year capped maximum age look better than the full max age. 

```{r residual plots}
d20 <- d20 |> mutate(facet = "Maximum age 20 yrs")
d30 <- d30 |> mutate(facet = "Maximum age 30 yrs")
dfull <- dfull |> mutate(facet = "Full age range")

res30plot <- ggplot(d30, aes(log_age_rel, res)) +
  geom_point(size = 0.5, alpha = 0.5) +
  geom_hline(yintercept = 0, color = "tomato",linewidth = 0.7)+
  labs(x = "Relative age (log)", y = "Residuals")+
  facet_wrap(~facet, strip.position = "right")+
  theme_bw(base_size = 8)+
  theme( panel.grid = element_blank(),
    axis.title.x = element_text(size = 7),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.y = element_text(size = 7),
    panel.border = element_rect(linewidth = 0.3),
     strip.background = element_rect(fill = "white", colour = NA, linewidth = 0.3),
    strip.text = element_text(size = 6, face = "plain"))

res20plot <- ggplot(d20, aes(log_age_rel, res)) +
  geom_point( size = 0.5, alpha = 0.5) +
  geom_hline(yintercept = 0, color = "tomato",linewidth = 0.7)+
  labs(x = "Relative age (log)", y = "Residuals")+
   facet_wrap(~facet, strip.position = "right")+
  theme_bw(base_size = 8)+
  theme( panel.grid = element_blank(),
    axis.title.x = element_text(size = 7),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.y = element_text(size = 7),
    panel.border = element_rect(linewidth = 0.3),
     strip.background = element_rect(fill = "white", colour = NA, linewidth = 0.3),
    strip.text = element_text(size = 6, face = "plain"))


resfullplot <- ggplot(dfull, aes(log_age_rel, res)) +
  geom_point( size = 0.5, alpha = 0.5) +
  geom_hline(yintercept = 0, color = "tomato",linewidth = 0.7)+
  labs(x = "Relative age (log)", y = "Residuals")+
   facet_wrap(~facet, strip.position = "right")+
  theme_bw(base_size = 8)+
  theme( panel.grid = element_blank(),
    axis.title.x = element_text(size = 7),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.y = element_text(size = 7),
    panel.border = element_rect(linewidth = 0.3),
     strip.background = element_rect(fill = "white", colour = NA, linewidth = 0.3),
    strip.text = element_text(size = 6, face = "plain"))

resplots <- (res20plot/res30plot/resfullplot)+plot_layout(axis_titles = "collect")+
  plot_annotation(tag_levels = "A")

# ggsave(
#  filename = "residuals.png",
#  plot = resplots,
#  dpi = 600,
#  path = "03_graphics/",
#  width = 8,
#  height = 12,
#  units = "cm")

```

### compare model coefficients

```{r compare model coefs}
coef_compare <- bind_rows(
  tidy(m23, effects = "fixed") |> 
    mutate(model = "Maximum age 20 yrs")) |> 
  filter(!term %in% c("(Intercept)" , "log_age_rel")) |> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST", 
                          term == "log_age_rel" ~ "Age")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST","Age" )))

cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","grey30","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2" )

colorspace::swatchplot(cbPalette, cvd = TRUE)

comp <- ggplot(coef_compare, aes(x = estimate, y = term)) +
  geom_point(
    size = 1.2) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    width = 0.1,
    linewidth = 0.3
  ) +
  geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  theme_classic(base_size = 9) +
  theme(
    panel.grid = element_blank(),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank()
  )
comp

# ggsave(
#  filename = "coeff_compare.png",
#  plot = comp,
#  dpi = 600,
#  path = "03_graphics/",
#  width = 8,
#  height = 6,
#  units = "cm")
```

### plot growth at age with different warming scenarios

```{r plot how model predicts growth at age}
nd_age <- expand_grid(
  sst_space_sc = 0,
  sst_time_sc = 0,
  log_age_rel = seq(min(d20$log_age_rel), max(d20$log_age_rel), length.out = 50)
) |>
  mutate(
    FishID = d$FishID[1],
    species = d$species[1],
    cohort = d$cohort[1]
  )

nd_age_pos <- expand_grid(
  sst_space_sc = 0,
  sst_time_sc = 2,
  log_age_rel = seq(min(d20$log_age_rel), max(d20$log_age_rel), length.out = 50)
) |>
  mutate(
    FishID = d$FishID[1],
    species = d$species[1],
    cohort = d$cohort[1]
  )

nd_age_neg <- expand_grid(
  sst_space_sc = 0,
  sst_time_sc = -2,
  log_age_rel = seq(min(d20$log_age_rel), max(d20$log_age_rel), length.out = 50)
) |>
  mutate(
    FishID = d$FishID[1],
    species = d$species[1],
    cohort = d$cohort[1]
  )

# predict again with linear model on this new data
p_age <- predict(m23, newdata = nd_age, re_form_iid = NA, se_fit = TRUE)
p_age_pos <- predict(m23, newdata = nd_age_pos, re_form_iid = NA, se_fit = TRUE)
p_age_neg <- predict(m23, newdata = nd_age_neg, re_form_iid = NA, se_fit = TRUE)


pplot <- ggplot(d20, aes(x = log_age_rel, y = log_increment_sc)) +
  geom_point(alpha = 0.1, size = 0.7) +
  geom_line(
    data = p_age,
    aes(x = log_age_rel, y = est, colour = "0"),
    inherit.aes = FALSE,
    linewidth = 0.7
  ) +
  geom_line(
    data = p_age_pos,
    aes(x = log_age_rel, y = est, colour = "+"),
    inherit.aes = FALSE,
    linewidth = 0.7
  ) +
  geom_line(
    data = p_age_neg,
    aes(x = log_age_rel, y = est, colour = "-"),
    inherit.aes = FALSE,
    linewidth = 0.7
  ) +
  scale_colour_manual(
    values =c("0" = "grey35", "+" = "#D55E00", "-" = "#0072B2")
  ) +
  guides(colour = guide_legend(ncol = 3)) +
  labs(x = "Relative age (log)", y = "Increment width (zscore)", colour = "SST")+
  theme_bw(base_size = 8)+
  theme( panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(hjust = 0.5, size = 6),
    legend.text = element_text(hjust = 0, size = 7),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 7),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.y = element_text(size = 7),
    panel.border = element_rect(linewidth = 0.3))

pplot

# ggsave(
#  filename = "modelled_growth_at_age.png",
#  plot = pplot,
#  dpi = 600,
#  path = "03_graphics/",
#  width = 8,
#  height = 8,
#  units = "cm")

```

### full model plots

```{r full model}
# plot full linear model with all interactions

nd <- expand_grid(
  sst_space_sc = c(
    quantile(d20$sst_space_sc, probs = 0.1),
    quantile(d20$sst_space_sc, probs = 0.9)
  ),
  sst_time_sc = modelr::seq_range(d20$sst_time_sc, n = 50),
  log_age_rel = c(
    quantile(d20$log_age_rel, probs = 0.1),
    quantile(d20$log_age_rel, probs = 0.5),
    quantile(d20$log_age_rel, probs = 0.9)
  ),
) |>
  mutate(
    FishID = d20$FishID[1],
    species = d20$species[1],
    cohort = d20$cohort[1]
  )

pred_m23 <- predict(
  m23,
  newdata = nd,
  re_form_iid = NA,
  se_fit = TRUE
)

pred_m23 <- pred_m23 |>
  mutate(
    age_group = factor(
      round(log_age_rel, 2),
      levels = sort(unique(round(log_age_rel, 2))),
      labels = c("Youngest 10%", "Median age", "Oldest 10%")
    ),
    sst_space_lab = factor(
      round(sst_space_sc, 1),
      labels = c("Coldest 10%", "Warmest 10%")
    )
  )

mean_sst_time <- mean(d20$sst_time, na.rm = TRUE)
sd_sst_time   <- sd(d20$sst_time, na.rm = TRUE)

pal <- c("steelblue","#B2182B")

m23_plot <- ggplot(pred_m23,
  aes(x = sst_time_sc,
    y = est,
    colour = sst_space_lab,
    fill = sst_space_lab)) +
  geom_ribbon(aes(
      ymin = est - 1.96 * est_se,
      ymax = est + 1.96 * est_se),
    alpha = 0.18,
    colour = NA) +
  geom_line(linewidth = 0.6) +
  facet_wrap(~ age_group, nrow = 1) +
  scale_colour_manual(values = c("Coldest 10%" = "steelblue", "Warmest 10%" = "#B2182B")) +
  scale_fill_manual(values = c("Coldest 10%" = "steelblue", "Warmest 10%" = "#B2182B")) +
  guides(fill = "none",
      colour = guide_legend(
      ncol = 2,
      title.position = "top",
      title.hjust = 0.5)) +
  labs(colour = "Mean site temperature",
    x = expression("Temporal SST anomaly (" * degree * "C)"),
    y = "Predicted increment width (zscore)") +
  scale_x_continuous(
    breaks = seq(-3, 3, by = 2),
    labels = round( seq(-3, 3, by = 2) * sd_sst_time + mean_sst_time, 1)) +
  theme_bw(base_size = 8) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "centre",
    legend.box.just = "centre",
    legend.title = element_text(size = 7, hjust = 0.5),
    legend.text = element_text(size = 6),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 5),
    axis.ticks = element_line(linewidth = 0.3),
    panel.border = element_rect(linewidth = 0.4),
    panel.spacing.x = unit(0.5, "lines"),
    strip.background = element_rect(fill = "white", colour = NA, linewidth = 0.3),
    strip.text = element_text(size = 6, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 6))
  )

m23_plot

# ggsave(
#  filename = "main_figure.png",
#  plot = m23_plot,
#  dpi = 600,
#  path = "03_graphics/",
#  width = 8,
#  height = 7,
#  units = "cm")

pchange <- pred_m23 |>
  group_by(log_age_rel, sst_space_lab) |>
  summarise(
    est_start = est[which.min(sst_time_sc)],
    est_end   = est[which.max(sst_time_sc)],
    change_est = est_end - est_start,
    pct_change_from_start = ((est_end - est_start) / abs(est_start)) * 100,
    .groups = "drop"
  )

```

Temporal SST anomaly was standardised prior to modelling; axis labels are shown here in the original °C units for interpretability.

```{r TSR plot}
df <- d20 |> group_by(site) |> 
  summarise(sst_space_sc = median(sst_space_sc))

nd_tsr <-
  expand_grid(
    # sst_space_sc = 0,
    sst_space_sc = c(-1, 0, 2),
    sst_time_sc = modelr::seq_range(d20$sst_time_sc, n = 50),
    log_age_rel = c(
    quantile(d20$log_age_rel, probs = 0.1),
    quantile(d20$log_age_rel, probs = 0.9)
  ),
  ) |>
  mutate(
    FishID = d20$FishID[1],
    species = d20$species[1],
    cohort = d20$cohort[1]
  )



pred_m23_tsr <- predict(
  m23,
  newdata = nd_tsr,
 re_form_iid = NA,
  se_fit = TRUE
)


age_vals <- sort(unique(round(pred_m23_tsr$log_age_rel, 2)))
space_vals <- sort(unique(round(pred_m23_tsr$sst_space_sc, 2)))

pred_m23_tsr <- pred_m23_tsr |>
  mutate(
    age_group = factor(
      round(log_age_rel, 2),
      levels = age_vals,
      labels = c("Youngest 10%", "Oldest 10%")
    ),
    space_group = factor(
      round(sst_space_sc, 2),
      levels = space_vals,
      labels = c("Cool sites", "Average sites", "Warm sites")
    )
  )

mean_sst_time <- mean(d20$sst_time, na.rm = TRUE)
sd_sst_time   <- sd(d20$sst_time, na.rm = TRUE)

cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","grey30","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2" )
colorspace::swatchplot(cbPalette, cvd = TRUE)

tsr_plot <- ggplot(
  pred_m24_tsr,
  aes(
    x = sst_time_sc,
    y = est,
    colour = age_group,
    fill = age_group)) +
  geom_ribbon(
    aes(
      ymin = est - 1.96 * est_se,
      ymax = est + 1.96 * est_se),
    alpha = 0.18,
    colour = NA) +
  geom_line(linewidth = 0.6) +
  facet_wrap(~ space_group, nrow = 1) +
  scale_colour_manual(
    values = c(
      "Youngest 10%" = cbPalette[3],
      "Oldest 10%" = cbPalette[4])) +
  scale_fill_manual(
    values = c(
      "Youngest 10%" = cbPalette[3],
      "Oldest 10%" = cbPalette[4]) ) +
  scale_x_continuous(
    breaks = seq(-3, 3, by = 2),
    labels = round( seq(-3, 3, by = 2) * sd_sst_time + mean_sst_time, 1)) +
  guides(
    fill = "none",
    colour = guide_legend(
      ncol = 2,
      title.position = "top",
      title.hjust = 0.5)) +
  labs(
    colour = "Relative age",
    x = expression("Temporal SST anomaly (" * degree * "C)"),
    y = "Predicted increment width (zscore)" ) +
  theme_bw(base_size = 8) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "centre",
    legend.box.just = "centre",
    legend.title = element_text(size = 7, hjust = 0.5),
    legend.text = element_text(size = 6),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 5),
    axis.ticks = element_line(linewidth = 0.3),
    panel.border = element_rect(linewidth = 0.4),
    panel.spacing.x = unit(0.5, "lines"),
    strip.background = element_rect(fill = "white", colour = NA, linewidth = 0.3),
    strip.text = element_text(size = 6, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 6)))

tsr_plot

# ggsave(
#  filename = "tsr_figure.png",
#  plot = tsr_plot,
#  dpi = 600,
#  path = "03_graphics/",
#  width = 8,
#  height = 7,
#  units = "cm")
```


```{r TSR plot map}

pred_m23_tsr_v <- predict(
  m23,
  newdata = nd_tsr,
 re_form_iid = NA,
  se_fit = TRUE
)

age_vals <- sort(unique(round(pred_m23_tsr_v$log_age_rel, 2)))
space_vals <- sort(unique(round(pred_m23_tsr_v$sst_space_sc, 2)))

pred_m23_tsr_v <- pred_m23_tsr_v |>
  mutate(
    age_group = factor(
      round(log_age_rel, 2),
      levels = age_vals,
      labels = c("Youngest 10%", "Oldest 10%")
    ),
    space_group = factor(
      round(sst_space_sc, 2),
      levels = space_vals,
      labels = c("Cold sites" , "Average sites","Warm sites")
    )
  ) |> 
  mutate(space_group = factor(space_group, ordered = TRUE, levels = c("Warm sites","Average sites","Cold sites" )))

cool_base <- pred_m23_tsr_v |> 
  filter(space_group == "Cold sites")

cool_slopes <- bind_rows(
  cool_base |>  mutate(space_group = "Average sites"),
  cool_base |>  mutate(space_group = "Warm sites")) |> 
  mutate(
    space_group = factor(
      space_group, ordered = TRUE,
     levels = c("Warm sites","Average sites","Cold sites" )) )

tsr_plot_v <- ggplot(
  pred_m23_tsr_v,
  aes(
    x = sst_time_sc,
    y = est,
    colour = age_group,
    fill = age_group)) +
  geom_ribbon(
    aes(
      ymin = est - 1.96 * est_se,
      ymax = est + 1.96 * est_se),
    alpha = 0.18,
    colour = NA) +
  geom_line(linewidth = 0.6) +
   geom_line(
    data = cool_slopes,
    aes(x = sst_time_sc, y = est, colour = age_group),
    linetype = "dashed",
    linewidth = 0.4,
    alpha = 0.55,
    inherit.aes = FALSE) +
  facet_wrap(~ space_group, ncol = 1, strip.position = "right") +
  scale_colour_manual(
  values = c("Youngest 10%" = cbPalette[3], "Oldest 10%" = cbPalette[4])) +
  scale_fill_manual(
  values = c("Youngest 10%" = cbPalette[3], "Oldest 10%" = cbPalette[4]))+
  scale_x_continuous(
    breaks = seq(-3, 3, by = 2),
    labels = round( seq(-3, 3, by = 2) * sd_sst_time + mean_sst_time, 1)) +
  guides(
    fill = "none",
    colour = guide_legend(
      ncol = 2,
      title.position = "top",
      title.hjust = 0.5) ) +
  labs(
    colour = "Relative age",
    x = expression("Temporal SST anomaly (" * degree * "C)"),
    y = "Predicted increment width (zscore)") +
  theme_bw(base_size = 8) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "centre",
    legend.box.just = "centre",
    legend.title = element_text(size = 7, hjust = 0.5),
    legend.text = element_text(size = 6),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 5),
    axis.ticks = element_line(linewidth = 0.3),
    panel.border = element_rect(linewidth = 0.4),
    #panel.spacing.x = unit(0.5, "lines"),
    panel.spacing.y = unit(0.35, "lines"),
    strip.background = element_rect(fill = "white", colour = NA, linewidth = 0.3),
    strip.text = element_text(size = 6, face = "plain"),
    axis.title.x = element_text(margin = margin(t = 6)),
    panel.background = element_rect(fill = NA, colour = NA),
        plot.background  = element_rect(fill = NA, colour = NA)
  )

tsr_plot_v

map_plot2 <- map_plot +
  theme(plot.margin = margin(2, -2, 2, 2))

tsr_plot_v <- tsr_plot_v +
  theme(plot.margin = margin(2, 2, 2, -2))

fig_side <- plot_grid(
  map_plot2,
  tsr_plot_v,
  nrow = 1,
  rel_widths = c(1.25, 0.75),
  align = "h",
  axis = "tb",
  labels = c("A", "B"),
  label_size = 10,
  label_fontface = "plain",
  label_x = c(0.02, 0.01),
  label_y = c(0.98, 0.98),
  hjust = 0,
  vjust = 1
)

fig_side



# ggsave(
#   "map_tsr_side_by_side.png",
#   fig_side,
#   path = "03_graphics/",
#   width = 8,
#   height = 12,
#   units = "cm",
#   dpi = 600, 
#   bg = "white"
# )
# 
# ggsave(
#   "tsr_plot.png",
#   tsr_plot_v,
#   path = "03_graphics/",
#   device = ragg::agg_png,
#   width = 3,
#   height = 8,
#   units = "cm",
#   dpi = 600,
#   bg = "transparent"
# )

```
