---
title: "sensitivity testing"
author: "Sarah Willington"
date: "2026-03-10"
output: html_document
---

Sensitivity testing for Willington et al. 2026 otolith paper

1. compare models with relative age to a maximum of 20 or 30 years

2. Randomly removing species to test for a strong species effect

3. Randomly remove ~ 20% of fish 

4. Sequentially remove 1 species at a time and refit the model 

5. Remove top 3 species by number of individuals 

6. compare to a weighted model

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


```

```{r load libraries}
library(tidyverse)
library(here)
library(patchwork)
library(effects) #fit and visualise the model
library(ggeffects) #view partial effects
library(tidylog)
library(mgcv)
library(sdmTMB)
library(modelr)

```

## load data and scale variables

```{r load data}

project_dir <- here::here()

d <- read_csv(file.path(project_dir, "0_data", "02_cleaned_data", "otolith_analysis.csv"))

d <- d |>
  mutate(
    FishID = as.factor(FishID),
    cohort = as.factor(cohort),
    species = as.factor(species),
    species_cohort = as.factor(species_cohort),
    species_site_cohort = as.factor(species_site_cohort), 
    site_cohort = as.factor(site_cohort)
  ) 

d20 <- d |> 
  filter(age<=20) |> 
  mutate(age_rel = if_else(maximum_age > 20, age / 20, age / maximum_age),
         age_sc = (age - mean(age)) / sd(age),
         log_age_sc = (log_age - mean(log_age)) / sd(log_age),
         log_increment_sc = (log_increment - mean(log_increment)) / sd(log_increment),
         .by = species) |> 
  mutate(log_age_rel = log(age_rel))

d30 <- d |> 
  filter(age<=30) |> 
  mutate(age_rel = if_else(maximum_age > 30, age / 30, age / maximum_age),
         age_sc = (age - mean(age)) / sd(age),
         log_age_sc = (log_age - mean(log_age)) / sd(log_age),
         log_increment_sc = (log_increment - mean(log_increment)) / sd(log_increment),
         .by = species)|> 
  mutate(log_age_rel = log(age_rel))

dfull <- d |> 
  mutate(age_rel = age/maximum_age,
         age_sc = (age - mean(age)) / sd(age),
         log_age_sc = (log_age - mean(log_age)) / sd(log_age),
         log_increment_sc = (log_increment - mean(log_increment)) / sd(log_increment),
         .by = species)|> 
  mutate(log_age_rel = log(age_rel))

```

## 1. fit full models

Compare models with age relative to a maximum of 20 years and age relative to a maximum of 30 years


```{r compare full models}
m23 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d20, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m33 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d30, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)


coef_compare <- bind_rows(
  tidy(m23, effects = "fixed") |> 
    mutate(model = "Maximum age 20 yrs"),
  tidy(m33, effects = "fixed") |> 
    mutate(model = "Maximum age 30 yrs")) |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
         model = factor(model, ordered = TRUE, levels = c("Maximum age 20 yrs", "Maximum age 30 yrs")))

cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","grey30","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2" )

comp <- ggplot(coef_compare, aes(x = estimate, y = term, colour = model)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_point(
    position = position_dodge(width = 0.45, reverse = TRUE),
    size = 1
  ) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.45, reverse = TRUE),
    width = 0.16,
    linewidth = 0.3
  ) +
  scale_color_manual(values = c(cbPalette[1], cbPalette[2])) +
  guides(colour = guide_legend(ncol = 2)) +
  labs(
    x = "Estimate",
    y = NULL,
    colour = "Model"
  ) +
  theme_classic(base_size = 9) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 5),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank())
comp

 ggsave(
  filename = "fullmodel_compare.png",
  plot = comp,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 6,
  units = "cm")
```

## 2. randomly remove species

Test out the species effect by randomly sampling the data to remove 3 or 6 species and refit the model

```{r remove 3 species}
set.seed(124)

collapse_removed <- function(x) paste(sort(unique(x)), collapse = "; ")

random_species_removal <- function(df, dataset_name, n_remove, n_draws = 50) {
  
  map_dfr(seq_len(n_draws), function(i) {
    
    # 1) sample species to remove
    species_to_remove <- df %>%
      distinct(species_id, species) %>%
      slice_sample(n = n_remove)
    
    # 2) remove them
    df_removed <- df %>%
      anti_join(species_to_remove, by = c("species_id", "species"))
    
    # 3) store removal info
    removed_ids   <- collapse_removed(species_to_remove$species_id)
    removed_names <- collapse_removed(species_to_remove$species)
    
    # 4) fit model
    fit <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = df_removed, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)
    
    # 5) tidy coefs + add metadata
    tidy(fit) %>%
      mutate(
        draw = i,
        n_removed = n_remove,
        df = dataset_name,
        group = paste(dataset_name, paste0(n_remove, " species"), sep = " - "),
        removed_species_id = removed_ids,
        removed_species = removed_names
      )
  })
}

results_all <- bind_rows(
  random_species_removal(df = d20, dataset_name = "Max 20", n_remove = 3, n_draws = 50),
  random_species_removal(df = d20, dataset_name = "Max 20", n_remove = 6, n_draws = 50),
  random_species_removal(df = d30, dataset_name = "Max 30", n_remove = 3, n_draws = 50),
  random_species_removal(df = d30, dataset_name = "Max 30", n_remove = 6, n_draws = 50)
)



coef_df <- results_all |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
         group = factor(group, ordered = TRUE, levels = c("Max 30 - 6 species",
                                                          "Max 30 - 3 species",
                                                          "Max 20 - 6 species",
                                                          "Max 20 - 3 species")),
          dataset = factor(df, ordered = TRUE, levels = c("Max 20", "Max 30")),
          removal = factor(n_removed, ordered = TRUE, levels = c(3, 6),
                           labels = c("3 species", "6 species")))

#write_csv(coef_df, "02_pipeline/02_sens_testing/final_plots/random_species_coeffs.csv")

  
cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2", "grey30")
colorspace::swatchplot(cbPalette, cvd = TRUE)

ran_sp <- ggplot(coef_df, aes(x = estimate, y = term, fill = group)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_boxplot(position = position_dodge(width = 0.8), 
               width = 0.6, 
               linewidth = 0.1, 
               outlier.size = 0.02) +
 scale_fill_manual(
    values = c(
      "Max 20 - 3 species" = cbPalette[1],
      "Max 20 - 6 species" = cbPalette[3],
      "Max 30 - 3 species" = cbPalette[2],
      "Max 30 - 6 species" = cbPalette[4]),
    guide = guide_legend(reverse = TRUE, nrow = 2, byrow = TRUE))+
  labs(
    x = "Estimate",
    y = NULL, 
    fill = NULL) +
  theme_classic(base_size = 9) +
  theme(
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 5),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank())

ran_sp


ggsave(
  filename = "species_removed.png",
  plot = ran_sp,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 8,
  units = "cm")



```

## 3. randomly remove fish 

randomly remove n fish from the data and refit the model 


```{r random sample}

set.seed(124)

nfish <- d |> 
    distinct(FishID) # ~20% of 722 = 144

collapse_removed <- function(x) paste(sort(unique(x)), collapse = "; ")

random_fish_removal <- function(df, dataset_name, n_remove, n_draws = 50) {
  
  map_dfr(seq_len(n_draws), function(i) {
    
    # 1) sample 100 FishID to remove (unique IDs)
  fish_to_remove <- df |> 
    distinct(FishID) |> 
    slice_sample(n = n_remove )  

  # 2) remove them 
  df_removed <- df |> 
    anti_join(fish_to_remove, by = "FishID")

  removed_fish <- collapse_removed(fish_to_remove$FishID)
    
    # 4) fit model
    fit <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = df_removed, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)
    
    # 5) tidy coefs + add metadata
    tidy(fit) %>%
      mutate(
        draw = i,
        n_removed = n_remove,
        df = dataset_name,
        group = paste(dataset_name, paste0(n_remove, " fish"), sep = " - "),
        removed_FishID = removed_fish
      )
  })
}

results_all_fish <- bind_rows(
  random_fish_removal(df = d20, dataset_name = "Max 20", n_remove = 144, n_draws = 50),
  random_fish_removal(df = d30, dataset_name = "Max 30", n_remove = 144, n_draws = 50)
)


coef_df_fish <- results_all_fish |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
          dataset = factor(df, ordered = TRUE, levels = c("Max 20", "Max 30"), 
                           labels = c("Maximum age 30 yrs", "Maximum age 20 yrs")))

#write_csv(coef_df_fish, "02_pipeline/02_sens_testing/final_plots/random_fish_coeffs.csv")

  
cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2", "grey30")
colorspace::swatchplot(cbPalette, cvd = TRUE)

ran_fish <- ggplot(coef_df_fish, aes(x = estimate, y = term, fill = dataset)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_boxplot(position = position_dodge(width = 0.8), 
               width = 0.5, 
               linewidth = 0.1, 
               outlier.size = 0.02) +
 scale_fill_manual(
    values = c(
     cbPalette[2],
     cbPalette[1]),
    guide = guide_legend(reverse = TRUE))+
  labs(
    x = "Estimate",
    y = NULL, 
    fill = NULL) +
  theme_classic(base_size = 9) +
  theme(
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 5),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank())

ran_fish


ggsave(
  filename = "fish_removed.png",
  plot = ran_fish,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 6,
  units = "cm")

```

## 4. remove 1 species 

Sequentially remove 1 species at a time to see if any one species has a large influence

```{r sequentially remove species}

systematic_species_removal <- function(df, dataset_name) {
  
  species_list <- df %>%
    distinct(species_id, species) %>%
    arrange(species_id)
  
  map_dfr(seq_len(nrow(species_list)), function(i) {
    
    # 1) choose species to remove
    species_to_remove <- species_list[i, ]
    
    # 2) remove it
    df_removed <- df %>%
      anti_join(species_to_remove, by = c("species_id", "species"))
    
    
    # 4) fit model
     fit <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = df_removed, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)
    
    # 5) tidy coefficients + metadata
    tidy(fit) %>%
      mutate(
        draw = i,
        dataset = dataset_name,
        removed_species_id = species_to_remove$species_id,
      removed_species = species_to_remove$species
      )
  })
}

results_all_sys <- bind_rows(
  systematic_species_removal(df = d20, dataset_name = "Max 20"),
  systematic_species_removal(df = d30, dataset_name = "Max 30")
)

coef_df_sys <- results_all_sys |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
          dataset = factor(dataset, ordered = TRUE, levels = c("Max 20", "Max 30"), 
                           labels = c("Maximum age 30 yrs", "Maximum age 20 yrs")))

#write_csv(coef_df_sys, "02_pipeline/02_sens_testing/final_plots/sys_species_coeffs.csv")

  
cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2", "grey30")
colorspace::swatchplot(cbPalette, cvd = TRUE)

sys_sp <- ggplot(coef_df_sys, aes(x = estimate, y = term, fill = dataset)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_boxplot(position = position_dodge(width = 0.8), 
               width = 0.5, 
               linewidth = 0.1, 
               outlier.size = 0.02) +
 scale_fill_manual(
    values = c(
     cbPalette[2],
     cbPalette[1]),
    guide = guide_legend(reverse = TRUE))+
  labs(
    x = "Estimate",
    y = NULL, 
    fill = NULL) +
  theme_classic(base_size = 9) +
  theme(
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 5),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 5),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 3.5),
    axis.ticks.y = element_blank())

sys_sp


ggsave(
  filename = "sys_species_removed.png",
  plot = sys_sp,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 6,
  units = "cm")
```

## 5. remove big 3 species 

 remove the 3 species with the most individuals 
 
```{r}

species_remove <- d %>%
  filter(species_id =="CSP"| species_id == "NFU" | species_id == "PBA") |> 
  distinct(species_id, species) %>%
  arrange(species_id)

  d20_removed <- d20 %>%
    anti_join(species_remove, by = c("species_id", "species"))
  d30_removed <- d30 %>%
    anti_join(species_remove, by = c("species_id", "species"))
  
removed_20 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d20_removed, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

removed_30 <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d30_removed, 
  family = student(link = "identity", df = NULL), 
  reml = TRUE)


coef_compare_b3 <- bind_rows(
  tidy(m23, effects = "fixed") |> 
    mutate(model = "Max 20 - all species"),
  tidy(removed_20, effects = "fixed") |> 
    mutate(model = "Max 20 - 3 species removed"),
  tidy(m33, effects = "fixed") |> 
    mutate(model = "Max 30 - all species"),
  tidy(removed_30, effects = "fixed") |> 
    mutate(model = "Max 30 - 3 species removed")) |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
         model = factor(model, ordered = TRUE, levels = c("Max 20 - all species", 
                                                          "Max 20 - 3 species removed",
                                                          "Max 30 - all species",
                                                          "Max 30 - 3 species removed")))


cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2", "grey30")

b3_plot <- ggplot(coef_compare_b3, aes(x = estimate, y = term, colour = model)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_point(
    position = position_dodge(width = 0.45, reverse = TRUE),
    size = 0.8
  ) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.45, reverse = TRUE),
    width = 0.16,
    linewidth = 0.25
  ) +
  scale_color_manual( values = c(
      "Max 20 - all species" = cbPalette[1],
      "Max 20 - 3 species removed" = cbPalette[3],
      "Max 30 - all species" = cbPalette[2],
      "Max 30 - 3 species removed" = cbPalette[4]),
    guide = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    x = "Estimate",
    y = NULL,
    colour = "Model"
  ) +
  theme_classic(base_size = 9) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 5),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank())

ggsave(
  filename = "big3.png",
  plot = b3_plot,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 8,
  units = "cm")


```

## 6. weigthed model

weight the model by the inverse of the number of individuals per species

```{r weighted model}

d20_w <- d20 |> 
  group_by(species) |> 
  mutate(
    n_fish_species = n_distinct(FishID),
    weight_fish = 1 / n_fish_species) |> 
  ungroup() |> 
  mutate(
    weight_fish = weight_fish / mean(weight_fish))

d30_w <- d30 |> 
  group_by(species) |> 
  mutate(
    n_fish_species = n_distinct(FishID),
    weight_fish = 1 / n_fish_species) |> 
  ungroup() |> 
  mutate(
    weight_fish = weight_fish / mean(weight_fish))


m23_weighted <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d20_w,
  weights = d20_w$weight_fish,
  family = student(link = "identity", df = NULL), 
  reml = TRUE)

m33_weighted <- sdmTMB(
  log_increment_sc ~ log_age_rel * sst_space_sc + 
    log_age_rel * sst_time_sc + 
    sst_space_sc * sst_time_sc +
    (1 | FishID) + (1 | species) + (1 | site_cohort),
  spatial = "off",
  data = d30_w,
  weights = d30_w$weight_fish,
  family = student(link = "identity", df = NULL), 
  reml = TRUE)


coef_compare_weight <- bind_rows(
  tidy(m23, effects = "fixed") |> 
    mutate(model = "Max 20 unweighted"),
  tidy(m23_weighted, effects = "fixed") |> 
    mutate(model = "Max 20 weighted"),
  tidy(m33, effects = "fixed") |> 
    mutate(model = "Max 30 unweighted"),
  tidy(m33_weighted, effects = "fixed") |> 
    mutate(model = "Max 30 weighted")) |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
         model = factor(model, ordered = TRUE, levels = c("Max 20 unweighted", 
                                                          "Max 20 weighted",
                                                          "Max 30 unweighted",
                                                          "Max 30 weighted")))


cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2", "grey30")

weighted_plot <- ggplot(coef_compare_weight, aes(x = estimate, y = term, colour = model)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_point(
    position = position_dodge(width = 0.55, reverse = TRUE),
    size = 0.8
  ) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.55, reverse = TRUE),
    width = 0.16,
    linewidth = 0.25
  ) +
  scale_color_manual( values = c(
      "Max 20 unweighted" = cbPalette[1],
      "Max 20 weighted" = cbPalette[3],
      "Max 30 unweighted" = cbPalette[2],
      "Max 30 weighted" = cbPalette[4]),
    guide = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    x = "Estimate",
    y = NULL,
    colour = "Model"
  ) +
  theme_classic(base_size = 9) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 5),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank())

weighted_plot


ggsave(
  filename = "weighted.png",
  plot = weighted_plot,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 8,
  units = "cm")


```

```{r 3 way comparison}
coef_compare_weight2 <- bind_rows(
  tidy(m23, effects = "fixed") |> 
    mutate(model = "Max 20 unweighted"),
  tidy(m23_weighted, effects = "fixed") |> 
    mutate(model = "Max 20 weighted"),
  tidy(m33, effects = "fixed") |> 
    mutate(model = "Max 30 unweighted"),
  tidy(m33_weighted, effects = "fixed") |> 
    mutate(model = "Max 30 weighted"),
  tidy(removed_20, effects = "fixed") |> 
    mutate(model = "Max 20 - 3 species"),
  tidy(removed_30, effects = "fixed") |> 
    mutate(model = "Max 30 - 3 species")) |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
         model = factor(model, ordered = TRUE, levels = c("Max 20 unweighted", 
                                                          "Max 20 weighted",
                                                          "Max 20 - 3 species",
                                                          "Max 30 unweighted",
                                                          "Max 30 weighted",
                                                          "Max 30 - 3 species")))


cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2", "grey30")

weighted_plot2 <- ggplot(coef_compare_weight2, aes(x = estimate, y = term, colour = model)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_point(
    position = position_dodge(width = 0.7, reverse = TRUE),
    size = 0.8
  ) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.7, reverse = TRUE),
    width = 0.16,
    linewidth = 0.2
  ) +
  scale_color_manual( values = c(
      "Max 20 unweighted" = cbPalette[1],
      "Max 20 weighted" = cbPalette[3],
      "Max 20 - 3 species" = cbPalette[6],
      "Max 30 unweighted" = cbPalette[2],
      "Max 30 weighted" = cbPalette[4],
      "Max 30 - 3 species" = cbPalette[7]),
    guide = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    x = "Estimate",
    y = NULL,
    colour = "Model"
  ) +
  theme_classic(base_size = 9) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 3),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank())

weighted_plot2


ggsave(
  filename = "weighted2.png",
  plot = weighted_plot2,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 8,
  units = "cm")
```

```{r 3 way comp shapes}
coef_compare_weight3 <- bind_rows(
  tidy(m23, effects = "fixed") |> 
    mutate(model = "Maximum age 20 yrs", 
           test = "Unweighted"),
  tidy(m23_weighted, effects = "fixed") |> 
    mutate(model = "Maximum age 20 yrs", 
           test = "Weighted"),
  tidy(m33, effects = "fixed") |> 
    mutate(model = "Maximum age 30 yrs", 
           test = "Unweighted"),
  tidy(m33_weighted, effects = "fixed") |> 
    mutate(model = "Maximum age 30 yrs", 
           test = "Weighted"),
  tidy(removed_20, effects = "fixed") |> 
    mutate(model = "Maximum age 20 yrs", 
           test = "3 species removed"),
  tidy(removed_30, effects = "fixed") |> 
    mutate(model = "Maximum age 30 yrs", 
           test = "3 species removed")) |> 
  filter(!term %in% c("(Intercept)", "log_age_rel"))|> 
  mutate(term = case_when(
                          term == "sst_space_sc" ~ "Spatial SST", 
                          term == "sst_time_sc" ~ "Temporal SST", 
                          term == "log_age_rel:sst_space_sc" ~ "Relative age x Spatial SST", 
                          term == "log_age_rel:sst_time_sc" ~ "Relative age x Temporal SST", 
                          term == "sst_space_sc:sst_time_sc" ~ "Spatial SST x Temporal SST")) |> 
  mutate(term = factor(term, ordered = TRUE, levels = c("Spatial SST x Temporal SST",
                                                        "Relative age x Temporal SST",
                                                        "Temporal SST",
                                                        "Relative age x Spatial SST",
                                                        "Spatial SST" )),
         model = factor(model, ordered = TRUE, levels = c("Maximum age 20 yrs", 
                                                          "Maximum age 30 yrs")), 
         test = factor(test, ordered = TRUE, levels = c("Unweighted", "Weighted", "3 species removed")))


cbPalette <- c("#56B4E9","#CC79A7", "#E69F00","#EED202", "#000000",
                "#D55E00","#009E73", "#0072B2", "grey30")

weighted_plot3 <- ggplot(coef_compare_weight3, 
                         aes(x = estimate, y = term, colour = model, shape = test)) +
   geom_vline(xintercept = 0, linetype = 2, linewidth = 0.3, colour = "grey60") +
  geom_point(
    position = position_dodge(width = 0.7, reverse = TRUE),
    size = 0.8
  ) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.7, reverse = TRUE),
    width = 0.4,
    linewidth = 0.2
  ) +
  scale_color_manual( values = c(
      "Maximum age 20 yrs" = cbPalette[1],
     "Maximum age 30 yrs"= cbPalette[2]),
    guide = guide_legend(ncol = 1, byrow = TRUE)) +
  labs(
    x = "Estimate",
    y = NULL,
    colour = "Model"
  ) +
  theme_classic(base_size = 9) +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
  legend.box = "vertical",
  legend.justification = "left",
  legend.box.just = "left",
  legend.spacing.y = unit(-0.1, "cm"),
    legend.title = element_blank(),
    legend.text = element_text(hjust = 0, size = 4.5),
    legend.key.height = unit(0.35, "cm"),
    legend.key.width = unit(0.5, "cm"),
    axis.title.x = element_text(size = 6),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.ticks.y = element_blank())

weighted_plot3


ggsave(
  filename = "weighted3.png",
  plot = weighted_plot3,
  dpi = 600,
  path = "02_pipeline/02_sens_testing/final_plots",
  width = 8,
  height = 8.5,
  units = "cm")
```

